\documentclass[addpoints]{exam}
\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\definecolor{light-gray}{gray}{0.95}
\newcommand{\code}[1]{\colorbox{light-gray}{\texttt{#1}}}
\usepackage{listings}
\usepackage{xcolor}

\usepackage{hyperref}
\usepackage{tikz,lipsum,lmodern}
\usepackage[most]{tcolorbox}
\usepackage{url}


%This block of commented code translates default words to Spanish
%-------------------------------------------------------------
%\pointpoints{punto}{puntos}
%\bonuspointpoints{punto extra}{puntos extra}

%\totalformat{Pregunta \thequestion: \totalpoints{} puntos}

%\chqword{Pregunta}
%\chpgword{Página}
%\chpword{Puntos}
%\chbpword{Puntos extra}
%\chsword{Puntos obtenidos}
%\chtword{Total}

%\boxedpoints
%-------------------------------------------------------------

\begin{document}
%This code creates the text before the first question
%-------------------------------------------------------------------
\begin{center}
\fbox{\fbox{\parbox{5.5in}{\centering
Experimenting with Community Structure}}}
\end{center}

% \vspace{5mm}
% \begin{tcolorbox}[colback=black!5!white,colframe=white!75!black]
% You can do the exercises in the order that you prefer. \\  
% the \textit{Going further} experiments take more time and I do not expect everyone to do them.
% \end{tcolorbox}
% \vspace{5mm}

%Here, the questions begin
\begin{questions}



\question Detecting your first Community Structure

\begin{tcolorbox}[colback=black!5!white,colframe=white!75!black]
To detect communities, you can use the \code{cdlib} package. It also contains functions for evaluation and comparison of partitions. For details, check the documentation at \code{\url{https://cdlib.readthedocs.io/en/latest/}}

\textbf{!!!} With google colab, you can install it with \code{!pip install cdlib}, but, due to a colab bug, you also need to do first: 

\code{!pip uninstall python-louvain}
validate with \code{y}, then
\code{!pip install python-louvain}.
\end{tcolorbox}
\begin{parts}
\part Using \code{networkx}, load the  airport dataset, provided as a graphml file. (reminder: you can download it in colab with \code{!wget URL} with URL the url of the file.
\part Using \code{cdlib}, detect communities on this network using the louvain method. You have to use the \code{algorithms.louvain} method (and do \code{from cdlib import algorithms} before.
%\part Add the communities found on your graph as a node property, using \code{NodeCluster.to\_node\_community\_map()} and \code{nx.set\_node\_attributes}. 
\part Visualize the communities found. In order to interpret them, you should draw each node at its geographical location, with a color per community. 

\begin{tcolorbox}[colback=black!5!white,colframe=white!75!black]
There are several ways to draw a spatial network with colors corresponding to communities, from using Gephi to plotting points on an interactive map using \code{folium}.

Here, I provide a simple code to plot the data as a simple scatter plot

\begin{lstlisting}[language=Python, caption=plot on a map]
import seaborn as sns
import matplotlib.pyplot as plt
x= list(nx.get_node_attributes(g,"lon").values())
y= list(nx.get_node_attributes(g,"lat").values())

coms_dict=coms.to_node_community_map()
hues=list(coms_dict[n][0] for n in g.nodes())

plt.figure(figsize=(12,8))
sns.scatterplot(x=x,y=y,hue=hues,palette=sns.color_palette
("tab20",len(coms.communities)),s=5)
\end{lstlisting}
\end{tcolorbox}

\part Vary the resolution parameter and observe changes in the community structure.

\end{parts}    




\vspace{5mm}
\question Comparing Partitions
%This question has several parts
\begin{parts}
\part The provided airport data also contains information about the country of each airport, which can be interpreted as a \textit{ground truth} partition of the network. 

You can obtain it using \code{nx.get\_node\_attributes(g,"country")}. Transform this information into a \code{NodeClustering} object of \code{cdlib}

(\code{nc = NodeClustering(partition,graph,"GroundTruth")}, with \code{partition} a list of list of nodes.
\part Compute the AMI, NMI, or another similarity score (\url{https://cdlib.readthedocs.io/en/latest/reference/evaluation.html}) between the community structure and the partition in countries
\part By exploring with a for loop the values of the resolution parameter for modularity, find the partition with the highest similarity to the partition in countries.
\part Compare visually the results, and try to interpret it. Is the partition in country a meaningful topological partition, i.e., is studying this network by considering that nodes in a same country form a coherent/homogeneous group a good approach? (yes and no, probably...)
\end{parts}





\vspace{5mm}
\question Going further : Intuitions on the SBM
\begin{tcolorbox}[colback=black!5!white,colframe=white!75!black]
I propose this exercise using only networkx and cdlib. You could do much more with SBM using \code{graph-tool} package (real SBM inference, degree-corrected SBM, Hierarchical SBM, etc.), but it requires a little bit more time to get used to at first, so I recommend it only if you're particularly interested in the topic.
\end{tcolorbox}
%This question has several parts
\begin{parts}
\part Compute the block matrix for a reasonable partition of the graph. For a given partition, you need to count the number of edges between and inside each community.
\part Using networkx \code{stochastic\_block\_model} method, generates a graph based on the computed block matrix
\part Using the network and node descriptors that you know, compare the properties of this generated graph with the properties of your original graph (and with a simple ER or configuration model). How is it different? How is it similar? Think about clustering coefficient, average distance, distribution of node centralities, degree distribution, etc.
\part How do these properties change when you increase/decrease the number of blocks?
\end{parts}




\end{questions}


\end{document}